Place of death in NRS COVID-19 deaths: visualisation for 2022 data

### 2022-06-09 covid_nrs_place_of_death_visualisations_2022.Rmd
finished the home deaths increase vs cumulative covid rate by LA plot

@@

@@

```{r}
(figure_map_home_deaths_increase_from_2020_la <-
  merged_deaths_la_historic_period_2015_2019 %>%
  filter(historic_period == "2015-2019") %>%
  group_by(ref_area, place_of_death) %>%
  summarise(
    ratio_deaths_to_historic = sum(deaths_all_causes, na.rm=TRUE) / sum(deaths_mean, na.rm=TRUE),
    percent_increase = ratio_deaths_to_historic - 1,
    .groups = "drop"
  ) %>%
  filter(place_of_death == "Home & other non-institution") %>%
  left_join(shapefile_la, by=c("ref_area"="LAD20NM")) %>%
  ggplot(aes(geometry=geometry, fill=percent_increase), colour="grey10") +
  geom_sf(colour="grey10") +
  # scale_fill_viridis_b(limits = c(minimum_ratio_between_hb_and_la_home_deaths-1, maximum_ratio_between_hb_and_la_home_deaths-1), labels = function(x) scales::percent(x, accuracy = 1)) +
  scale_fill_viridis_b(limits = c(0, 0.6), labels = function(x) scales::percent(x, accuracy = 1)) +
  theme(axis.text = element_blank(), panel.grid = element_blank()) +
  labs(x="", y="",
       title = str_wrap("Increased home deaths since the start of 2020 relative to historical mean 2015-2019, by Local Authority", width=40),
       # subtitle = str_wrap("A ratio of 1 means there was no increase; A ratio of 1.6 means there was a 60% increase of home deaths in 2020 over the mean 2015-2019.", width = 80),
       # caption = paste0(c(areas_relatively_higher_number_home_deaths_text, home_deaths_ratio_text, source_nrs_text), collapse="\n"),
       caption = paste0(c(source_nrs_text), collapse="\n"),
       fill = "Home death\nincrease relative\nto 2015-2019"
       ) +
  NULL)

@@

```{r}
(figure_map_home_deaths_increase_by_year_la <-
  merged_deaths_la_historic_period_2015_2019 %>%
  filter(historic_period == "2015-2019") %>%
  group_by(year, ref_area, place_of_death) %>%
  summarise(
    ratio_deaths_to_historic = sum(deaths_all_causes, na.rm=TRUE) / sum(deaths_mean, na.rm=TRUE),
    percent_increase = ratio_deaths_to_historic - 1,
    .groups = "drop"
  ) %>%
  filter(place_of_death == "Home & other non-institution") %>%
  left_join(shapefile_la, by=c("ref_area"="LAD20NM")) %>%
  ggplot(aes(geometry=geometry, fill=percent_increase), colour="grey10") +
  geom_sf(colour="grey10") +
  # scale_fill_viridis_b(limits = c(minimum_ratio_between_hb_and_la_home_deaths-1, maximum_ratio_between_hb_and_la_home_deaths-1), labels = function(x) scales::percent(x, accuracy = 1)) +
  scale_fill_viridis_b(limits = c(0, 0.6), labels = function(x) scales::percent(x, accuracy = 1)) +
  theme(
    axis.text = element_blank(), 
    panel.grid = element_blank(),
    # legend.position = "bottom"
    ) +
  facet_wrap(~year, nrow = 1) +
  labs(x="", y="",
       title = "Increased home deaths since the start of 2020 relative to historical mean 2015-2019, by Local Authority",
       caption = paste0(c(source_nrs_text), collapse="\n"),
       fill = "Home death\nincrease relative\nto 2015-2019"
       ) +
  NULL)

@@

```{r}
# TODO: continue working on this!
(figure_map_covid_cases_from_2020_la <-
  weekly_covid_cases_la %>%
  group_by(ref_area) %>%
  summarise(
    ratio_covid_cases = max(crude_rate_positive_per100k),
    .groups = "drop"
  ) %>% 
  # left_join(population_estimate_la_2019, by = c("ref_area"="la")) %>%
  # dplyr::group_by(ref_area) %>%
  # dplyr::summarise(
  #   ratio_covid_cases = sum(weekly_positive, na.rm=TRUE) / all_ages *1e5, 
  #   .groups = "drop"
  # ) %>%
  left_join(shapefile_la, by=c("ref_area"="LAD20NM")) %>%
  ggplot(aes(geometry=geometry, fill=ratio_covid_cases), colour="grey10") +
  geom_sf(colour="grey10") +
  scale_fill_viridis_b(option = "B", labels = function(x) format(x, big.mark=",")) +
  theme(axis.text = element_blank(), panel.grid = element_blank()) +
  labs(x="", y="",
       title = str_wrap("Total COVID-19 cases since the beginning of 2020 per 100,000 population, by Local Authority", width=50),
       subtitle = "",
       caption = paste0(c(source_phs_text), collapse="\n"),
       fill = str_wrap("Cases per 100,000 population", width = 15)
       ) +
  NULL)

figure_map_covid_cases_from_2020_la %>%
save_output_file(filename = file.path(dir_outputs, paste0(c("figure_map_covid_cases_from_2020_la", latest_data_modified_date), collapse = "_")), extensions = c(".pdf",".png"), plot = ., width = 13, height = 14, dpi = 300, units = "cm")
```

>>>>

crude_rate_text <- "Total cases = crude cumulative positive cases over time."

@@

## Re-merge weekly deaths with 2015-2019 data for 2022 for LA & HB

@@

```{r}
(figure_map_home_deaths_increase_from_2020_la <-
  merged_deaths_la_historic_period_2015_2019 %>%
  filter(historic_period == "2015-2019") %>%
  group_by(ref_area, place_of_death) %>%
  summarise(
    ratio_deaths_to_historic = sum(deaths_all_causes, na.rm=TRUE) / sum(deaths_mean, na.rm=TRUE),
    percent_increase = ratio_deaths_to_historic - 1,
    .groups = "drop"
  ) %>%
  filter(place_of_death == "Home & other non-institution") %>%
  left_join(shapefile_la, by=c("ref_area"="LAD20NM")) %>%
  ggplot(aes(geometry=geometry, fill=percent_increase), colour="grey10") +
  geom_sf(colour="grey10") +
  # scale_fill_viridis_b(limits = c(minimum_ratio_between_hb_and_la_home_deaths-1, maximum_ratio_between_hb_and_la_home_deaths-1), labels = function(x) scales::percent(x, accuracy = 1)) +
  scale_fill_viridis_b(limits = c(0, 0.6), labels = function(x) scales::percent(x, accuracy = 1)) +
  theme(axis.text = element_blank(), panel.grid = element_blank()) +
  labs(x="", y="",
       title = str_wrap("Increased home deaths since the start of 2020 relative to historical mean 2015-2019, by Local Authority", width=40),
       # subtitle = str_wrap("A ratio of 1 means there was no increase; A ratio of 1.6 means there was a 60% increase of home deaths in 2020 over the mean 2015-2019.", width = 80),
       # caption = paste0(c(areas_relatively_higher_number_home_deaths_text, home_deaths_ratio_text, source_nrs_text), collapse="\n"),
       caption = paste0(c(only_2022_text,source_nrs_text), collapse="\n"),
       fill = "Home death\nincrease relative\nto 2015-2019"
       ) +
  NULL)

@@

```{r}
(figure_map_home_deaths_increase_by_year_la <-
  merged_deaths_la_historic_period_2015_2019 %>%
  filter(historic_period == "2015-2019") %>%
  group_by(year, ref_area, place_of_death) %>%
  summarise(
    ratio_deaths_to_historic = sum(deaths_all_causes, na.rm=TRUE) / sum(deaths_mean, na.rm=TRUE),
    percent_increase = ratio_deaths_to_historic - 1,
    .groups = "drop"
  ) %>%
  filter(place_of_death == "Home & other non-institution") %>%
  left_join(shapefile_la, by=c("ref_area"="LAD20NM")) %>%
  ggplot(aes(geometry=geometry, fill=percent_increase), colour="grey10") +
  geom_sf(colour="grey10") +
  # scale_fill_viridis_b(limits = c(minimum_ratio_between_hb_and_la_home_deaths-1, maximum_ratio_between_hb_and_la_home_deaths-1), labels = function(x) scales::percent(x, accuracy = 1)) +
  scale_fill_viridis_b(limits = c(0, 0.6), labels = function(x) scales::percent(x, accuracy = 1)) +
  theme(
    axis.text = element_blank(), 
    panel.grid = element_blank(),
    # legend.position = "bottom"
    ) +
  facet_wrap(~year, nrow = 1) +
  labs(x="", y="",
       title = "Increased home deaths since the start of 2020 relative to historical mean 2015-2019, by Local Authority",
       caption = paste0(c(only_2022_text,source_nrs_text), collapse="\n"),
       fill = "Home death\nincrease relative\nto 2015-2019"
       ) +
  NULL)

@@

```{r}
(figure_map_covid_cases_from_2020_la <-
  weekly_covid_cases_la %>%
  group_by(ref_area) %>%
  summarise(
    ratio_covid_cases = max(crude_rate_positive_per100k),
    .groups = "drop"
  ) %>% 
  left_join(shapefile_la, by=c("ref_area"="LAD20NM")) %>%
  ggplot(aes(geometry=geometry, fill=ratio_covid_cases), colour="grey10") +
  geom_sf(colour="grey10") +
  scale_fill_viridis_b(option = "B", labels = function(x) format(x, big.mark=",")) +
  theme(axis.text = element_blank(), panel.grid = element_blank()) +
  labs(x="", y="",
       title = str_wrap("Total COVID-19 cases since the start of 2020, by Local Authority", width=40),
       subtitle = "",
       caption = paste0(c(only_2022_text,crude_rate_text,source_phs_text), collapse="\n"),
       fill = str_wrap("Cases per 100,000 population", width = 15)
       ) +
  NULL)

figure_map_covid_cases_from_2020_la %>%
save_output_file(filename = file.path(dir_outputs, paste0(c("fig_map_covid_cases_from_2020_la", latest_data_modified_date), collapse = "_")), extensions = c(".pdf",".png"), plot = ., width = 13, height = 14, dpi = 300, units = "cm")

## combine maps
(figure_map_home_deaths_increase_vs_covid_case_from_2020_la <-
  wrap_plots(
    figure_map_home_deaths_increase_from_2020_la +
      theme(plot.margin = margin(0,0,0,0), legend.box.margin = margin(0,0,0,0))
      ,
    figure_map_covid_cases_from_2020_la +
      theme(plot.margin = margin(0,0,0,0), legend.box.margin = margin(0,0,0,0))
  )
)

figure_map_home_deaths_increase_vs_covid_case_from_2020_la %>%
save_output_file(filename = file.path(dir_outputs, paste0(c("fig_map_home_deaths_increase_vs_covid_case_from_2020_la", latest_data_modified_date), collapse = "_")), extensions = c(".pdf",".png"), plot = ., width = 24, height = 16, dpi = 300, units = "cm")
```

#### LA covid rates over time

I didn't finish this plot because the scale goes from 500 per 100k to 25k per 100k, which is a pain to depict in a meaningful way. Mayeb with a separate scale for each plot?

```{r}
## the problem with this plot is that there were between 500 per 100k cases to 25k per 100k cases over time and over LAs - how do we visualise that?
(figure_map_covid_cases_by_year_la <-
  weekly_covid_cases_la %>%
  left_join(population_estimate_la_2019 %>% rename(population_estimate = all_ages, year_estimate = year), by = c("ref_area"="la")) %>%
  group_by(ref_area, year, population_estimate) %>%
  summarise(
    cumulative_cases = sum(weekly_positive, na.rm=TRUE),
    ratio_covid_cases = cumulative_cases / population_estimate * 1e5,
    .groups = "drop"
  ) %>%
  distinct %>%
  left_join(shapefile_la, by=c("ref_area"="LAD20NM")) %>%
  # ggplot(aes(geometry=geometry, fill=log(ratio_covid_cases)), colour="grey10") +
  ggplot(aes(geometry=geometry, fill=ratio_covid_cases), colour="grey10") +
  geom_sf(colour="grey10") +
  # scale_fill_viridis_b(option = "B", labels = function(x) format(exp(x), big.mark=",")) +
  scale_fill_viridis_b(option = "B", labels = function(x) format(x, big.mark=",")) +
  theme(axis.text = element_blank(), panel.grid = element_blank()) +
  facet_wrap(~year, nrow = 1) +
  labs(x="", y="",
       title = str_wrap("Total COVID-19 cases since the start of 2020 per 100,000 population, by Local Authority", width=50),
       subtitle = "",
       caption = paste0(c(source_phs_text), collapse="\n"),
       fill = str_wrap("Cases per 100,000 population", width = 15)
       ) +
  NULL)
```
